Estimation of the perceived time of presence of sources in urban acoustic environment using deep learning techniques

### 2019-04-03 paper/paperDraft.tex
input for intro
Add cites + proofread

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{fancyhdr}
\usepackage{a4wide} % Smaller margins = more text per page.
\usepackage[hidelinks]{hyperref}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{pdflscape}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{bbm}
\usepackage{color}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\ml}[1]{\textcolor{red}{ML : #1}}
\pagestyle{fancy}
\begin{document}
\author{F\'elix Gontier $^1$, Catherine Lavandier $^2$, Pierre Aumond $^{2, 3}$\\Mathieu Lagrange $^1$, Jean-Francois Petiot $^1$}
\date{
$^1$\\
$^2$\\
$^3$
}
\title{Estimation of the perceived time of presence of sources in urban acoustic environment using deep learning techniques}
\maketitle

@@

\section{Introduction}
\label{sec:intro}
%- Du bruit au soundscape
%- La pleasantness premier axe sur les affects liés au soundscape
%- On a des modèles de pleasantness perceptifs qui font intervenir des temps de présence des sources
%- Peut-on à partir de la mesure physique faire des modèles de pleasantness sur la base d'indicateurs physiques?
%- Idée: si l'on fait de la reconnaissance de sources, on peut faire un meilleur modèle physique que les précédents
%- Première étape: On vérifie que le corpus créé est proche d'un corpus réel
%- Sur le corpus on fait de l'apprentissage profond
%- On compare le nouveau modèle physique deep en comparaison aux modèles physiques old school Cartazur (dimension purement niveau sonore) et Grafic (niveau sonore + Fréquence/sources)
In urban areas, the decrease of quality of sound environments resulting from urbanization is a major cause of annoyance, and has been identified as an important health factor. Consequently, the 2002/49/CE European directive requires the development of noise maps in large cities. These maps aim at providing useful tools for urban planning and noise reduction purposes. As an alternative to the noise-only paradigm used until recently, the soundscape approach defined in the ISO 12913-1 (CITE) norm as "the acoustic environment as perceived and understood and/or experienced by people and/or society, in context" extends the characterization of sound environments by considering the positive influence of some of its components.\\

The assessment of soundscape quality through standardized perceptual descriptors has been extensively studied (CITE). Specifically, the axis of pleasantness is increasingly associated with soundscape quality in the recent literature (CITE). Pleasantness is consistently modeled as a function of both the perceived sound level and the activity of sources of interest within urban environments (CITE). While the taxonomy used differs across studies, three major source types are found to contribute to pleasantness: traffic, human voices and birds. In existing models, source activity can be quantified by the dominance, emergence or time of presence.\\

Though, the perceptual characterization of soundscapes requires subjective inputs, and thus cannot directly be applied to the production of perceptually motivated noise maps. With the advent of the Internet of Things and the availability of low-cost sensor networks (CITE), it may be feasible to predict pleasantness or other perceptual factors from acoustic measurements.

Bridging the gap between raw data issued from the sensors to reasonable estimates of perceptual attributes is an interesting and ongoing area of research. Ricciardi et al. (CITE) present a physical model of pleasantness using only variations of the measured sound level. Though, because these indicators are not representative of the composition of the sound scene, the model underperforms in comparison to predictions using perceptual assessments. Aumond et al. (CITE) explore new indicators based on spectral variations in recordings of sound scenes. These indicators are designed to model the activity of specific sound sources, specifically human voices and birds, but their discriminative properties are limited.

In order to tackle this issue, we believe that the recent advances in machine learning can be useful to better approximate the contributions of sound sources relevant for perceptual attribute prediction such as pleasantness. In this paper, we do so by developing new indicators relying on source recognition models which are built based on deep learning techniques CITE. However, large amounts of annotated data are needed for these architectures to learn to extract relevant information from recordings and to the best of our knowledge such databases do not currently exist in the literature. In order to circumvent this issue, we propose in this paper to study the potential of the use of simulated corpora (CITE) \ml{j'enleverai c'est super ambigu, on a l'impression que l'on annote au hasard :) where sound scenes are composed from semi-randomly generated annotations}.

The contributions of this paper are thus three-fold:
\begin{enumerate}
  \item provide controled acoustic data and corresponding perceptual attributes relevant to the study of urban soundscapes perception
  \item propose numerical means for predicting time of presence of sound sources from raw acoustic data
  \item provide a comparative study of the behavior of the proposed approach with state of the art ones.
\end{enumerate}

First, a perceptual test is conducted where simulated sound scenes are compared to recordings to validate their use in the context of pleasantness prediction. Next, a deep learning architecture is designed and trained that performs multi-label sound source recognition at relevant time scales. Estimations of pleasantness using these predictions are next compared to existing models from physical indicators.

\ml{mettre les noms de section dans ce par}

\subsection{Perceptual validation}
\label{sec:pval}

\subsubsection{Sound scenes corpus}
\label{sec:pval_corp}

A first corpus is created to validate the use of simulated sound scenes in the development of pleasantness prediction models. It is based on recordings from one of the four soundwalks performed in~\cite{aumond2017}, including 19 locations in the 13th district of Paris. A classification of these scenes is proposed in~\cite{gloaguen2017} in terms of ambiance: \textit{park}, \textit{quiet street}, \textit{noisy street} and \textit{very noisy street}. For each of the 19 recordings 45 seconds of audio in a single channel are extracted that represent well the properties of their respective ambiances, without single events overwhelming their overall perception. The resulting 45~s extracts are replicated to study perceptual differences between recorded and simulated scenes. The recordings are first annotated in terms of the following properties:

@@

To further study the perceptual properties of simulated scenes, the corpus is extended with 75 simulated scenes corresponding to new scenarios. Because of the lack of large isolated sources datasets, we choose to restrict the scene composition to 3 sources: \textit{traffic}, \textit{human voices} and \textit{birds}. Considering the aim of the study, newly generated scenarios should cover most real-life situations while remaining perceptually plausible. Thus, statistics on background and event properties are extracted for each of the three sources from annotations presented in~\cite{gloaguen2017}. These statistics include, conditionally to the ambiance:

@@

Adjustments are made on the variance of the considered properties to better cover possible scenarios. The EBR of voice events, mostly read English recordings, are reduced to improve realism. An additional \textit{place} ambiance is added with properties derived empirically from available data. Diverse new scenarios are then generated by sampling the background and event properties distributions. For each scene the time of presence for the traffic, voice and bird sources is estimated using indicators proposed in~\cite{gontier2018}, and 75 scenes are chosen to maximize pairwise distance in the resulting 3-dimensional space. This is shown in Figure~\ref{fig:tvb_pres}. The playback sound level is also conditioned on the ambiance according to typical sound levels in urban context shown in Figure~\ref{fig:amb_levels}, and ranges from 46.6~dB to 77.1~dB.\\

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/lvls.jpg}
    \caption{Example of sound level scale a function of the environment for new simulated scenes. TODO CITE}\label{fig:amb_levels}
\end{figure}

\subsubsection{Experiment}
\label{sec:pval_exp}

@@

The scenes are played at a given sound level as discussed in Section~\ref{sec:pval_corp}, through the same computer and sound card configurations. Beyerdynamics DT-990 Pro headphones are used for all participants. These headphones are calibrated in a semi-anechoic chamber, resulting in a relation between playback Leq and LAeq and sound card output voltage for pink noise. This ensures that correctly scaled scenes will be heard at the desired sound level by listeners.\\

@@

\subsection{Indicators for pleasantness prediction}
\label{sec:inds}

Several indicators are computed from the audio signals that are found in previous studies~\cite{aumond2017, gontier2018, ricciardi2014} to correlate well with perceptual parameters.\\

@@

\subsection{Indicator prediction through deep learning}
\label{sec:deep}

\subsubsection{Learning corpus}
\label{sec:deep_corp}

@@

The generation procedure is the same for both datasets, and is equivalent to that of the simulated sub-corpus considered for the perceptual test in Section~\ref{sec:pval_corp}. Only traffic, human voices and birds sources compose the sound scenes due to the low number of available samples and recording annotations for other sources. The isolated samples database is split in the same proportions as the two datasets: two-thirds for the development set and the remaining one-third for the evaluation set. As a result, while the generation process is the same for both corpora both the scenarios and samples are different.\\

@@

Target outputs are associated to each input frame to train the model. Based on previous work the time of presence of traffic, voices and birds can be approximated by the $\hat T_s(\alpha, \beta)$ indicator. Thus, the source presence values before averaging along time can be used as weak labels. Alternatively, it is possible to perform sound level estimation using ground truth labels. Though, this regression task is difficult to learn compared to source detection, and will not be discussed here as it did not result in better time of presence estimations in our experiments.

\subsubsection{Model architecture}
\label{sec:deep_mdl}

@@

\subsection{Perceptual validation}
\label{sec:perc}
%=> 3.1, 3.2

\subsection{Baseline pleasantness models}
\label{sec:base}
%=> 3.3, 4.2

\subsubsection{Perceptual models}
\label{sec:base_perc}

@@

Table~\ref{tab:percm} summarizes performance metrics computed for the two models. $\hat P_{p1}$ and $\hat P_{p2}$ display similar estimation errors as well as no significant difference in adjusted $R^2$ and correlation to assessed pleasantness. The mean standard deviation of pleasantness assessments for this experiment is about 1.77, thus perceptual models are considered accurate with about 0.6 RMSE on the 10-point scale.

@@

\subsubsection{Physical models}
\label{sec:base_phys}

The analysis of physical indicators is done using the arithmetic mean of subjective assessments over all participants for replicated and simulated scenes. Two scenes containing only one source are excluded for implementation concerns, resulting in $n=92$ scenes.\\

@@

Subsequently, multilinear models of pleasantness are constructed based on the results of the analysis of correlations. As in Section~\ref{sec:base_perc} a VIF check is performed on predictors to ensure that no multi-collinearity exists. On the present corpus, the best model is:
\begin{equation}
\hat P_{\varphi 1} = 16.71 - 0.18~L_{50} + 1.01~\hat T_B(\alpha, \beta)
\end{equation}
Indicators related to traffic and voices activity are absent compared to the perceptual models in Section~\ref{sec:base_perc}. This is because the $L_{50}$ used as a predictor is correlated to both traffic parameters $L_{T, p}$ and $T_{T, p}$ that appeared in perceptual models $\hat P_{p1}$ and $\hat P_{p2}$ respectively, and no strong contribution of voices to pleasantness is identified in this corpus.\\

@@

\subsection{Prediction of pleasantness through deep learning}
\label{sec:pred}

Table~\ref{tab:perf_cmp} summarizes the performance of the source detection deep learning architecture presented in Section~\ref{sec:deep} on the evaluation dataset. The overall presence detection accuracy is 92.11\%. The model performs similarly well for the three sources, ranging from 90.08\% for birds to 94.09\% for traffic. Regarding the type of errors, they are equally split between false positives and false negatives. Though, these rates vary depending on the source. Traffic has a very low false positive rate at 2.46\% and high false negative rate, while birds display the highest false positive rate at 12.30\%. This is expected given the spectral content related to these sources. Traffic is usually located in lower frequencies than voice or bird events, but may contain high frequency components mistaken for birds by the model. The resulting error on the time of presence is 0.12 RMSE overall on a 0-1 scale, and is close for the three sources.\\

@@

Next, the capacity of these predictions to give relevant pleasantness estimates is studied. For this analysis, only the perceptual experiment is used as no assessments are available on the evaluation dataset. To evaluate the model's robustness to the increased polyphony and source complexity of scenes in real-life scenarios, the corpus is further split in two parts:
\begin{enumerate}
\item The 25 recorded and replicated scenes, which contain additional sources not present in the development or evaluation datasets simulated for the optimization of deep learning models,
\item The 75 simulated scenes that are obtained from the same simulation process as both the development and evaluation dataset.
\end{enumerate}
Pleasantness predictions are obtained by substituting time of presence estimates computed from the presence detection architecture's outputs to the $\hat P_{\varphi 1}$ model presented in Section~\ref{sec:base_phys}. Thus, only the outputs corresponding to the presence of birds are used. The results are compared to estimates using ground truth $\hat T_s(\alpha, \beta)$ labels in Table~\ref{tab:pppred}. First, pleasantness estimations are equally performant using the model's predictions or the ground truth labels , with an RMSE of about 0.84 and 82\% of explained variance on the perceptual experiment corpus. This performance of the detection model is expected given the low errors on time of presence estimates. Labels from the detection model result in higher errors on the first sub-corpus containing recorded and replicated scenes, with sources not seen during training. In both cases the error is still lower on the first sub-corpus, indicating that the detection model generalizes well to new content and sources.\\

@@

Pleasantness prediction using the detection model is as effective as the best baseline model from Section~\ref{sec:base_phys}, and outperforms previous models relying on physical indicators directly computed on the audio signal. Though, errors are significantly higher than the one of perceptual models in Section~\ref{sec:base_perc}.\\

@@

The perceptual models of pleasantness are comparable to those proposed in~\cite{ricciardi2014} and~\cite{aumond2017}, of respective expressions (translated from 1-11 scales to 0-10):
\begin{align}
\hat P_{p3} &= 7.11 - 0.38~OL - 0.15~T_{T, p} + 0.20~T_{V, p} + 0.15~T_{B, p}\\
\hat P_{p4} &= 8.70 - 0.47~OL - 0.21~T_{T, p} + 0.12~T_{V, p} + 0.09~T_{B, p}
\end{align}
Several hypotheses can be drawn from this result. First, the use of a simulated corpus with low source complexity does not alter the fundamental relationships between perceptual quantities, nor does the in vivo setting of the experiment. However, the latter may have an impact on the contribution of the overall loudness to pleasantness: the regression coefficient is $-0.67$, which is significantly higher than for in situ models. This may be due to scenes at the same sound level appearing louder when played through headphones in a laboratory setting.\\

@@

\bibliographystyle{plain}
\bibliography{refs}

>>>>

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{fancyhdr}
\usepackage{a4wide} % Smaller margins = more text per page.
\usepackage[hidelinks]{hyperref}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{pdflscape}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{bbm}
\usepackage{color}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\ml}[1]{\textcolor{red}{ML : #1}}
\pagestyle{fancy}
\begin{document}
\author{F\'elix Gontier $^1$, Catherine Lavandier $^2$, Pierre Aumond $^{2, 3}$\\Mathieu Lagrange $^1$, Jean-Francois Petiot $^1$}
\date{
$^1$ LS2N, UMR 6004, Ecole Centrale de Nantes, CNRS, F-44300, France\\
$^2$ ETIS, UMR 8051, University of Paris Seine, University of Cergy-Pontoise, ENSEA, CNRS, F-95000, France\\
$^3$ IFSTTAR, CEREMA, UMRAE, F-44344, Bouguenais, France
}
\title{Estimation of the perceived time of presence of sources in urban acoustic environment using deep learning techniques}
\maketitle

@@

\section{Introduction}
\label{sec:intro}

In urban areas, the decrease of quality of sound environments resulting from urbanization is a major cause of annoyance, and has been identified as an important health factor. Consequently, the 2002/49/CE European directive~\cite{ec2002} requires the development of noise maps in large cities. These maps aim at providing useful tools for urban planning and noise reduction purposes. As an alternative to the noise-only paradigm used until recently, the soundscape approach defined in the ISO 12913-1~\cite{iso2014} norm as "the acoustic environment as perceived and understood and/or experienced by people and/or society, in context" extends the characterization of sound environments by considering the positive influence of some of its components.\\

The assessment of soundscape quality through standardized perceptual descriptors has been extensively studied~\cite{viollon2000, axelsson2010, cain2013, jeon2018, aletta2016}. Specifically, the axis of pleasantness is found as the most explanatory of subjective assessments, and is thus increasingly associated with soundscape quality in the recent literature~\cite{axelsson2010, aumond2017, delaitre2014}. Pleasantness is consistently modeled as a function of both the perceived sound level and the activity of sources of interest within urban environments. While the taxonomy used differs across studies, three major source types are found to contribute to pleasantness in urban contexts: technological, human and natural sources~\cite{nilsson2007, axelsson2010}, which can be simplified as traffic, human voices and birds~\cite{lavandier2006, ricciardi2014, aumond2017}. In existing models, source activity can be quantified by the dominance, emergence or time of presence.\\

Though, the perceptual characterization of soundscapes requires subjective inputs, and thus cannot directly be applied to the production of perceptually motivated noise maps. With the advent of the Internet of Things and the availability of low-cost sensor networks~\cite{ardouin2018, mydlarz2017}, it may be feasible to predict pleasantness or other perceptual factors from acoustic measurements. Bridging the gap between raw data issued from the sensors to reasonable estimates of perceptual attributes is an interesting and ongoing area of research. Ricciardi et al.~\cite{ricciardi2014} present a physical model of pleasantness using only variations of the measured sound level. Though, because these indicators are not representative of the composition of the sound scene, the model underperforms in comparison to predictions using perceptual assessments. Aumond et al.~\cite{aumond2017} explore new indicators based on spectral variations in recordings of sound scenes. These indicators are designed to model the activity of specific sound sources, specifically human voices and birds, but their discriminative properties are limited.\\

In order to tackle this issue, we believe that the recent advances in machine learning can be useful to better approximate the contributions of sound sources relevant for perceptual attribute prediction such as pleasantness. In this paper, we do so by developing new indicators relying on source recognition models which are built based on deep learning techniques, which is a growing interest in the machine listening community~\cite{cakir2015, salamon2017-2}. However, large amounts of annotated data are needed for these architectures to learn to extract relevant information from recordings and to the best of our knowledge such databases do not currently exist in the literature. In order to circumvent this issue, we propose in this paper to study the potential of the use of simulated corpora~\cite{rossignol2015, salamon2017}.\\

The contributions of this paper are thus three-fold:
\begin{enumerate}
  \item Provide controlled acoustic data and corresponding perceptual attributes relevant to the study of urban soundscapes perception
  \item Propose numerical means for predicting the time of presence of sound sources from raw acoustic data
  \item Provide a comparative study of the behavior of the proposed approach with state of the art ones.
\end{enumerate}

First, a perceptual test is conducted in Section~\ref{sec:data} where simulated sound scenes are compared to recordings to validate their use in the context of pleasantness prediction. Next, Section~\ref{sec:methods} presents the design of a deep learning architecture that performs multi-label sound source recognition at relevant time scales. Estimations of pleasantness using these predictions are next compared to existing models from physical indicators in Section~\ref{sec:results}.

\section{Data}
\label{sec:data}

\subsection{Corpus generation}
\label{sec:data_corp}

A corpus is first created to validate the use of simulated sound scenes in the development of pleasantness prediction models. It is based on recordings from one of the four soundwalks performed in~\cite{aumond2017}, including 19 locations in the 13th district of Paris. A classification of these scenes is proposed in~\cite{gloaguen2017} in terms of ambiance: \textit{park}, \textit{quiet street}, \textit{noisy street} and \textit{very noisy street}. For each of the 19 recordings 45 seconds of audio in a single channel are extracted that represent well the properties of their respective ambiances, without single events overwhelming their overall perception. The resulting 45~s extracts are replicated to study perceptual differences between recorded and simulated scenes. The recordings are first annotated in terms of the following properties:

@@

To further study the perceptual properties of simulated scenes, the corpus is extended with 75 simulated scenes corresponding to new scenarios. Because of the lack of large isolated sources datasets, we choose to restrict the scene composition to 3 sources: \textit{traffic}, \textit{human voices} and \textit{birds}. Newly generated scenarios should cover most real-life situations while remaining perceptually plausible. Thus, statistics on background and event properties are extracted for each of the three sources from annotations presented in~\cite{gloaguen2017}. These statistics include, conditionally to the ambiance:

@@

Adjustments are made on the variance of the considered properties to better cover possible scenarios. The EBR of voice events, mostly read English recordings, are reduced to improve realism. An additional \textit{place} ambiance is added with properties derived empirically from available data. Diverse new scenarios are then generated by sampling the background and event properties distributions. For each scene the time of presence for the traffic, voice and bird sources is estimated using indicators proposed in~\cite{gontier2018}, and 75 scenes are chosen to maximize pairwise distance in the resulting 3-dimensional space. This is shown in Figure~\ref{fig:tvb_pres}. The playback sound level is also conditioned on the ambiance according to typical sound levels in urban context, and ranges from 46.6~dB to 77.1~dB.\\

@@

\subsection{Perceptual experiment}
\label{sec:data_exp}

@@

The scenes are played at a given sound level as discussed in Section~\ref{sec:data_corp}, through the same computer and sound card configurations. Beyerdynamics DT-990 Pro headphones are used for all participants. These headphones are calibrated in a semi-anechoic chamber, resulting in a relation between playback Leq and LAeq and sound card output voltage for pink noise. This ensures that correctly scaled scenes will be heard at the desired sound level by listeners.\\

@@

\subsection{Validation}
\label{sec:data_val}

\subsection{Indicators for pleasantness prediction}
\label{sec:methods_inds}

Based on previous studies~\cite{aumond2017, gontier2018, ricciardi2014}, several indicators are identified that correlate well with perceptual parameters. These indicators can be used to develop baseline physical models of pleasantness in the design of a deep learning architecture for sound source recognition.\\

@@

\subsection{Indicator prediction using deep learning}
\label{sec:methods_deep}

@@

The generation procedure is the same for both datasets, and is equivalent to that of the simulated sub-corpus considered for the perceptual test in Section~\ref{sec:data_corp}. Only traffic, human voices and birds sources compose the sound scenes due to the low number of available samples and recording annotations for other sources. The isolated samples database is split in the same proportions as the two datasets: two-thirds for the development set and the remaining one-third for the evaluation set. As a result, while the generation process is the same for both corpora both the scenarios and samples are different.\\

@@

Target outputs are associated to each input frame to train the model. In~\cite{gontier2018} the time of presence of traffic, voices and birds was successfully approximated by the $\hat T_s(\alpha, \beta)$ indicator. Thus, the source presence values obtained before averaging along time can be used as weak labels. Alternatively, it is possible to perform sound level estimation using ground truth labels. Though, this regression task is difficult to learn compared to source detection, and will not be discussed here as it did not result in better time of presence estimations in our experiments.\\

@@

\subsection{Perceptual pleasantness model}
\label{sec:results_perc}

@@

Table~\ref{tab:percm} summarizes performance metrics computed for the two models. $\hat P_{p1}$ and $\hat P_{p2}$ display similar estimation errors as well as no significant difference in adjusted $R^2$ and correlation to assessed pleasantness. The mean standard deviation of pleasantness assessments for this experiment is about 1.77 on the 11-point scale, the obtained perceptual models are thus considered accurate with about 0.6 RMSE.

@@

\subsection{Physical pleasantness models}
\label{sec:results_phys}

The analysis of physical indicators presented in Section~\ref{sec:methods_inds} is done using the arithmetic mean of subjective assessments over all participants for replicated and simulated scenes. Two scenes containing only one source are excluded for implementation concerns, resulting in $n=92$ scenes.\\

@@

Subsequently, multilinear models of pleasantness are constructed based on the results of the analysis of correlations. As in Section~\ref{sec:results_perc} a VIF check is performed on predictors to ensure that no multi-collinearity exists. On the present corpus, the best model is:
\begin{equation}
\hat P_{\varphi 1} = 16.71 - 0.18~L_{50} + 1.01~\hat T_B(\alpha, \beta)
\end{equation}
Indicators related to traffic and voices activity are absent compared to the perceptual models in Section~\ref{sec:results_perc}. This is because the $L_{50}$ used as a predictor is correlated to both traffic parameters $L_{T, p}$ and $T_{T, p}$ that appeared in perceptual models $\hat P_{p1}$ and $\hat P_{p2}$ respectively, and no strong contribution of voices to pleasantness is identified in this corpus.\\

@@

\subsection{Pleasantness prediction through deep learning}
\label{sec:results_deep}

The performance of the source detection deep learning architecture presented in Section~\ref{sec:methods_deep} on the evaluation dataset is summarized in Table~\ref{tab:perf_cmp}. The overall presence detection accuracy is 92.11\%. The model performs similarly well for the three sources, ranging from 90.08\% for birds to 94.09\% for traffic. Regarding the type of errors, they are equally split between false positives and false negatives. Though, these rates vary depending on the source. Traffic has a very low false positive rate at 2.46\% and high false negative rate, while birds display the highest false positive rate at 12.30\%. This is expected given the spectral content related to these sources. Traffic is usually located in lower frequencies than voice or bird events, but may contain high frequency components mistaken for birds by the model. The resulting error on the time of presence is 0.12 RMSE overall on a 0-1 scale, and is close for the three sources.\\

@@

Next, the capacity of these predictions to give relevant pleasantness estimates is studied. For this analysis, only the perceptual experiment is used as no assessments are available on the evaluation dataset. To evaluate the model's robustness to the increased polyphony and source complexity of scenes in real-life scenarios, the corpus is further split in two parts:
\begin{enumerate}
\item The 25 recorded and replicated scenes, which contain additional sources not present in the development or evaluation datasets simulated for the optimization of deep learning models,
\item The 75 simulated scenes that are obtained from the same simulation process as both the development and evaluation dataset.
\end{enumerate}
Pleasantness predictions are obtained by substituting time of presence estimates computed from the presence detection architecture's outputs to the $\hat P_{\varphi 1}$ model presented in Section~\ref{sec:results_phys}. Thus, only the outputs corresponding to the presence of birds are used. The results are compared to estimates using ground truth $\hat T_s(\alpha, \beta)$ labels in Table~\ref{tab:pppred}. First, pleasantness estimations are equally performant using the model's predictions or the ground truth labels , with an RMSE of about 0.84 and 82\% of explained variance on the perceptual experiment corpus. This performance of the detection model is expected given the low errors on time of presence estimates. Labels from the detection model result in higher errors on the first sub-corpus containing recorded and replicated scenes, with sources not seen during training. In both cases the error is still lower on the first sub-corpus, indicating that the detection model generalizes well to new content and sources.\\

@@

Pleasantness prediction using the detection model is as effective as the best baseline model from Section~\ref{sec:results_phys}, and outperforms previous models relying on physical indicators directly computed on the audio signal. Though, errors are significantly higher than the one of perceptual models in Section~\ref{sec:results_perc}.\\

@@

The perceptual models of pleasantness are comparable to those proposed in~\cite{ricciardi2014} and~\cite{aumond2017}, of respective expressions (translated from 1-11 scales to 0-10):
\begin{align}
\hat P_{p3} &= 7.11 - 0.38~OL - 0.15~T_{T, p} + 0.20~T_{V, p} + 0.15~T_{B, p}\\
\hat P_{p4} &= 8.70 - 0.47~OL - 0.21~T_{T, p} + 0.12~T_{V, p} + 0.09~T_{B, p}
\end{align}
Several hypotheses can be drawn from this result. First, the use of a simulated corpus with low source complexity does not alter the fundamental relationships between perceptual quantities, nor does the \textit{ex situ} setting of the experiment. However, the latter may have an impact on the contribution of the overall loudness to pleasantness: the regression coefficient is $-0.67$, which is significantly higher than for \textit{in situ} models. This may be due to scenes at the same sound level appearing louder when played through headphones in a laboratory setting.\\

@@

\bibliographystyle{unsrt}
\bibliography{refs}
